---
title: "progress-report-oriniker.Rmd"
author: "Olivia Riniker"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Lahman)
head(HallOfFame, 3)
View(HallOfFame)
library(rvest)
url <- "https://www.baseball-reference.com/awards/hof_2023.shtml"
html <- read_html(url)
tables <- html_table(html)
hof_2023 <- tables[[1]]
View(hof_2023)
```
 in my new scraped data set i'm going to have to go in a create and new variables like year id(just the year), playerID( which will be either a new ID i have to create or Ill have to each through the hall of fame data set and see this they already have a player ID), voted by( which will all be BBWAA), ballots(will be 389), needed(which is 292), votes(which is already in new data set), inducted (which will be yes id votes > needed), category( i believe they are all players), needed_notes( every entry in this column is NA so i believe these entries will be as well)
 
```{r}
library(rvest)
library(dplyr)

colnames(hof_2023) <- c("Rk", "Name", "YoB", "votes", "%vote", 
                        "HOFm", "HOFs", "Yrs", "WAR", "WAR7", "JAWS", "Jpos",
                        "G", "AB", "R", "H", "HR", "RBI", "SB", "BB", 
                        "BA", "OBP", "SLG", "OPS", "OPS+", "W", "L", 
                        "ERA", "ERA+", "WHIP", "G_pitch", "GS", "SV", 
                        "IP", "H_pitch", "HR_pitch", "BB_pitch", "SO_pitch", 
                        "Pos_summary")


hof_2023 <- hof_2023[-1, ] 

hof_2023$`%vote` <- readr::parse_number(hof_2023$`%vote`)


hof_2023$yearID <- 2023
  hof_2023$votedBy <- "BBWAA"
  hof_2023$ballots <- 389
  hof_2023$needed <- 292
  hof_2023$inducted <- ifelse(hof_2023$`%vote` >= 75.06, "Y", "N")
  hof_2023$category <- "Player"
  hof_2023$needed_note <- NA
# Remove any special characters or additional prefix from the names
hof_2023$Name <- gsub("^X-", "", hof_2023$Name)
hof_2023$Name <- gsub("á", "a", hof_2023$Name)
hof_2023$Name <- gsub("í", "i", hof_2023$Name)
hof_2023$Name <- gsub("J.J", "J. J", hof_2023$Name)
hof_2023$Name <- gsub("R.A", "R. A", hof_2023$Name)
hof_2023$Name <- gsub("HOF", "", hof_2023$Name)
hof_2023$Name <- sub("\\s+$", "", hof_2023$Name)

# Create full names in hof_2023
hof_2023$`First Lastname` <- hof_2023$Name

# Load the People dataset and find playerID
full_names <- People %>% mutate(
  `First Lastname`=paste(nameFirst, nameLast))
people_names <- People %>% select(playerID, `First Lastname`)

# Match playerID
hof_2023 <- merge(hof_2023, people_names, by = "First Lastname", all.x = TRUE)
  
  # Create new variables with fixed values


final_2023 <- hof_2023 %>% select(
  playerID, yearID, votedBy, 
  ballots, needed, votes, 
  inducted, category, needed_note 
  )
view(final_2023)
 
```
 
 After attempting to clean the data and shape it I found difficulties especially when trying to get the players ID. especially with player Francisco Rodriguez (because there are two in the People data set) as well as with Todd Helton because the HOF at the end messed with finding the player ID even after trying to get rid of and trailing white-space. 
 
 
 
 This last section is just concantonating my cleaned data set into the hall of fame data set and then creating a new HallOfFame csv into the repo.
 
```{r}
final_data <- rbind(HallOfFame, final_2023)

write.csv(final_data, "HallOfFame.csv", row.names=FALSE)


```


