---
title: "progress-report-oriniker.Rmd"
author: "Olivia Riniker"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Lahman)
head(HallOfFame, 3)
View(HallOfFame)
library(rvest)
url <- "https://www.baseball-reference.com/awards/hof_2023.shtml"
html <- read_html(url)
tables <- html_table(html)
hof_2023 <- tables[[1]]
View(hof_2023)
```
 in my new scraped data set i'm going to have to go in a create and new variables like year id(just the year), playerID( which will be either a new ID i have to create or Ill have to each through the hall of fame data set and see this they already have a player ID), voted by( which will all be BBWAA), ballots(will be 389), needed(which is 292), votes(which is already in new data set), inducted (which will be yes id votes > needed), category( i believe they are all players), needed_notes( every entry in this column is NA so i believe these entries will be as well)
 
```{r}
library(tidyverse)
hof_2023 <- hof_2023[-1, ]


#createing new names of the static variables
hof_2023$yearID <- 2023
hof_2023$votedBy <- "BBWAA"
hof_2023$ballots <- 389
hof_2023$needed <- 292
hof_2023$category <- "Player"
hof_2023$needed_note <- NA

#renaming to playerId

hof_2023[2] <- gsub("X-", "", hof_2023[2])

last_names <- sapply(strsplit(hof_2023[2], " "), tail, 1)


player_ID <- People %>% filter(nameLast == last_names)

hof_2023$playerID <- player_ID$playerID

hof_2023$playerID <- mapply(generate_playerID, hof_2023[,2])
head(hof_2023)



```
 
 
 
 
